---
title: "'Hail No!' Say Taxi Drivers to Uber"
author: "Cai Ytsma, Juliane Bombosch, Sofia Salcedo"
date: "May 10, 2018"
output: 
  html_document:
    code_folding: hide
    df_print: paged
---

***An Introduction to Uber***

A new epidemic has arisen, striking fear into the hearts of New York City Taxi drivers. This epidemic goes by many aliases, the most well known ones being "Uber" and "Lyft." Uber was officially launched in June 2010. Between 2011 and 2015, Uber has provided around 82,000 rides per day in New York City. By 2015, Uber was been named one of the most valuable startup companies in the world by Business Insider.
  
Using Uber data (2015) that was available, and comparing that to taxi data, we looked for general trends related to  domestic flights out of two airports: Newark and JFK. While the taxi and airport data was recorded in January 2017, a clear pattern in preference emerges.  

Ride sharing is so successful because it’s simple and marketed as the cheaper option. This is true about long trips, but not necessarily for short trips (94% of Uber trips) and surge pricing makes this option even more expensive in comparison to taxis [Time 2014](http://time.com/3556741/uber/). 
 
***So why is Uber the popular option?***

1. *Uber uses an app instead of physically hailing a cab.* For many people, this is more accessible- whether it’s because they are more comfortable with using technology or it removes the physical problem of hailing a cab (hailing a cab brings out both issues of disability as well as racial profiling on both the driver and passenger front).

2. *You can estimate the cost beforehand.* In taxis, there is little clarity before the ride about how expensive the ride will be. While it is an estimate, passengers can see whether they are getting a surge price, how long it will take, and this information is comforting to a lot of passengers. 

3. *It’s faster.* With the number of Uber drivers compared to taxi drivers, the wait for an Uber is shorter. And in a world where time is money, the number of minutes that you wait for a ride counts.

***How does this translate to New York City and airports?***

New York City has a lot of airports, which require a high number of taxis and ride-sharing to pick up the passengers. Two airports- Newark and John F. Kennedy can provide an example of the competition between Uber and taxis. JFK airport is 16 miles east of midtown Manhattan, located in Jamaica, Queens. Newark is 15 miles outside of New York City, 18 miles away from Midtown. This doesn’t appear to be a significant difference, but the time, and the fare dependent on that time, can be a big difference. 

Because of the fierce competition, other airports, like LAX in Los Angeles, have put restrictions on ride-sharing services that are not present for taxis in the same area. These restrictions include only being allowed to drive and pick up passengers in certain areas as well as not being allowed to wait for passengers. These restrictions make taxis the preferred method at LAX, unlike the rest of the city. Similarly, in 2016, Newark banned the use of Uber, preferring to support the licensed and regulated taxi drivers.

```{r, echo = FALSE, warning = FALSE, message = FALSE}
#calling the libraries
library(tidyverse)
library(mdsr)
library(RMySQL)
library(extrafont)

#define databases
nyctaxi <- dbConnect_scidb(dbname = "nyctaxi")
airlines <- dbConnect_scidb(dbname = "airlines")
uber_jan <- read.csv("./jan2015uber.csv")
taxi_rides <- read.csv("./taxi_rides.csv") %>%
  mutate(ride_type = "taxi")
```


```{r, echo = FALSE, warning = FALSE} 
#uber data
#jfk
uber_jfk_jan <- uber_jan %>%
  filter(locationID == 132) %>%
  group_by(day) %>%
  summarise(n_rides = n()) %>%
  mutate(airport = "JFK")

#newark
uber_ewr_jan <- uber_jan %>%
  filter(locationID == 1) %>%
  group_by(day) %>%
  summarise(n_rides = n()) %>%
  mutate(airport = "EWR")

#combine
uber_rides <- uber_jfk_jan %>%
  full_join(uber_ewr_jan, by = c("airport","day", "n_rides")) %>%
  mutate(ride_type = "uber")
```


```{r, label = "get flight info", warning = FALSE, echo = FALSE}
#flight info
#get flight data from MySQL
ny_flights_jan <- airlines %>%
  dbGetQuery("SELECT f.day, f.dest, p.seats FROM airlines.flights f
INNER JOIN airlines.planes p
	ON p.tailnum = f.tailnum
WHERE f.year = 2017
AND f.month = 1
AND f.dest in ('JFK', 'EWR');")

#estimate number of people leaving per day
people_per_day <- ny_flights_jan %>% 
  group_by(dest, day) %>%
  summarize(n_people = sum(seats), num_flights = n()) %>%
  ungroup() %>%
  mutate(flight_qual = n_people / 5, airport = dest) %>%
  select(-dest)
```


```{r, echo = FALSE, warning = FALSE}
#combining datasets
total_df <- people_per_day %>%
  full_join(uber_rides, by = c("day", "airport")) %>%
  full_join(taxi_rides, by = c("day", "airport", "n_rides", "ride_type"))
```

```{r, echo = FALSE}
#plot
ggplot(data = total_df, aes(x = day)) +
  geom_bar(aes(y = n_rides, fill = ride_type), stat = "identity") +
  scale_fill_manual("Ride Type", values = c("#ebac00", "black")) +
  geom_line(aes(y = flight_qual)) +
  theme(panel.grid = element_blank(),
        panel.background = element_blank(),
        text = element_text(family = "Calibri")) +
    scale_y_continuous("Number of Rides") +
    scale_x_continuous("Day") +
  facet_wrap(~airport, ncol = 1, labeller = as_labeller(c("EWR" = "Newark Liberty International Airport", "JFK" = "John F. Kennedy International Airport")))
```

<<<<<<< HEAD
***What does this all mean?***
Taxis are used at a higher rate compared to taxis at both airports, at such a great difference that you can barely see the Uber rides on the graph. 

There are enough restriction on Uber drivers that it is not as beneficial to them to pick up passengers at the airport. Showing the impending ban of Uber in 2016, Newark Uber data shows only 4 rides for the entire month! 

Aside from restrictions, someone who is at the airport, especially during the week (in January), is likely to be traveling for business. Those who are traveling for business on planes are more likely to be older. They are not the target demographic of Uber and would be less likely to trust the platform, relying on taxis and other forms of transportation to get back into the city. 

Uber is a popular service, but is it undercutting Uber in the way that is widely publicized? Not necessarily. The data shows that taxis have 98% of rides coming from the JFK and Newark airports. There are legislative protections that are being put into place to protect these staples in larger cities. There are still great percentages of people who rely on taxis as opposed to Ubers. At least at airports, taxis are still king.

It is important to note that data from multiple years was used. There was a restiction on what data was available because the data sets are so large. While general trends of competition between the two ride services can be measured, it is also important to note that it is not exact to compare these two data sets. 

```{r}
#comparing uber to taxi
=======
```{r, echo = FALSE, warning = FALSE}
#comparing number of uber to taxi rides
>>>>>>> 59f7f639c942b31b9bf1465bd18a2634803413c9
comparison <- total_df %>%
  select(n_rides, ride_type) %>%
  filter(!is.na(ride_type)) %>%
  group_by(ride_type) %>%
  summarize(N = sum(n_rides)) %>%
  ungroup() %>%
  mutate(prop = N / sum(N))
```

OLD STUFF:
calculated most average month
#popular_months <- airlines %>%
#  dbGetQuery("SELECT month, SUM(1) AS sum
#FROM airlines.flights
#WHERE dest = 'JFK'
#GROUP BY month
#ORDER BY sum DESC;")

#deviation <- popular_months %>%
#  summarize(mean = sum(sum)/12)

#mean <- deviation[[1]]

#most_avg_month <- popular_months %>%
#  mutate(near_avg = abs((sum/mean) - 1)) %>%
#  arrange(near_avg) %>%
#  head(1)

In trying to turn a string into just a day of a filtered month in the data package *nyctaxi yellow #2017* we couldn't figure out a way to select a string from **tpep_pickup_datetime** without losing all of the data atatched to it or to make a function that would be able to delete the first and last part of the string only leaving the date. The package was also so big that even with 50,000 rows we was not able to have all the days of the month included.
#SELECT * 
#FROM nyctaxi.yellow
#WHERE tpep_pickup_datetime LIKE  '%2017-03%';

#SELECT
#SUBSTRING_INDEX('2017-03-09 21:30:11', ' ' , -1)
#;
#SELECT
#SUBSTRING_INDEX('09 21:30:11', ' ' , 1)
#;
#SELECT
#SUBSTRING_INDEX('09', '0', -1)
#;
#SELECT RIGHT('2017-03-09 21:30:11', 11) 
#;
#SELECT LEFT('09 21:30:11', 2);
