---
title: "Mini Project 4"
author: "Cai Ytsma, Juliane Bombosch, Sofia Salcedo"
date: "April 27, 2018"
output: html_document
---
COMPARING APRIL 2005 and APRIL 2015

Using taxi and plane data, joining databases? To see connections between taxi and plane, in the shadow of ride sharing services.


```{r}
#calling the libraries
library(tidyverse)
library(mdsr)
library(RMySQL)
nyctaxi <- dbConnect_scidb(dbname = "nyctaxi")
airlines <- dbConnect_scidb(dbname = "airlines")
```

Sofia goals: gathering Uber and lyft info
1. create a database that has the dates of the information from Uber and lyft info
What am I trying to get out of that? 
- how many trips were made per day for each

- there's nothing in the tables that has lyft or uber data, the rows just say "null"

```{r}
https://raw.githubusercontent.com/fivethirtyeight/uber-tlc-foil-response/master/uber-trip-data/uber-raw-data-apr14.csv
```

for finding the location of the airport, try using PULocationID in the "yellow" taxi data

----------
Cai goals: estimate number of people leaving the airport by date

delineate by domestic/intl - find number cabs per capita and assume differences due to uber

```{r}
popular_months <- airlines %>%
  dbGetQuery("SELECT month, SUM(1) AS sum
FROM airlines.flights
WHERE dest = 'JFK'
GROUP BY month
ORDER BY sum DESC;")
```

```{r}
avg <- popular_months %>%
  summarize(mean = sum(sum)/12)

mean <- avg[[1]]

most_avg_month <- popular_months %>%
  mutate(near_avg = abs((sum/mean) - 1)) %>%
  arrange(near_avg) %>%
  head(1)
```



